### 2020-03-14 Common/chapters/abstract.tex
Julien and Johan: typo
Julien and Johan: minors

% !TEX root = ../main.tex
Automating the detection of cancer participates to an early detection and treatment, which increases the chances of recovery. Recent algorithms in artificial intelligence relying on deep learning have shown promising results in this field. Indeed, the usefulness of convolutional neural networks (CNNs) for segmentation or classification tasks is no longer to be proven. However, the performance of these models is often limited by the amount of data which is available to train the algorithm.

This thesis first presents a state-of-the-art convolutional neural network for pro\-state lesion classification. All steps from the data processing to the smallest detail regarding the neural network training are explained, ensuring a complete reproducibility of the experiment. This model is then evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, achieving an AUC of 0.76. This result constitutes a solid baseline and confirms the proper functioning of the implementation.

On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) is built. This method shows that integrating information from diverse datasets improves automated prostate cancer diagnosis. Indeed, it appears that cancerous lesions coming from various body parts share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier. This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancer can be taken advantage of. Thanks to this technique, the AUC achieved on our test set increases by 18\% (from 0.68 to 0.80).

>>>>

% !TEX root = ../main.tex
Automating the detection of cancer contributes to an early detection and treatment, which increases the chances of recovery. Recent algorithms in artificial intelligence relying on deep learning have shown promising results in this field. Indeed, the usefulness of convolutional neural networks (CNNs) for segmentation or classification tasks is no longer to be proven. However, the performance of these models is often limited by the amount of data which is available to train the algorithm.

This thesis first presents a state-of-the-art convolutional neural network for pro\-state lesion classification. All the steps from the data processing to the smallest detail regarding the neural network training are explained, ensuring a complete reproducibility of the experiment. This model is then evaluated on the official SPIE-AAPM-NCI Prostate MR Classification Challenge dataset, achieving an AUC of 0.76. This result constitutes a solid baseline and confirms the proper functioning of the implementation.

On top of this implementation, a new transfer learning approach using lesions of multiple body parts (brain and lung) is built. This method shows that integrating information from diverse datasets improves automated prostate cancer diagnosis. Indeed, it appears that cancerous lesions coming from various body parts share low-level features that can be used to increase the generalization ability and performance of the prostate lesion classifier. This technique provides a concrete solution to the lack of available data for prostate classification and suggests that many other types of cancers can be taken advantage of. Thanks to this technique, the AUC achieved on our test set increases by 18\% (from 0.68 to 0.80).

### 2020-03-14 Common/chapters/deep_learning.tex
Julien and Johan: minors

\chapter{Deep learning}
\label{ch:deep_learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work. It starts with a historical timeline of deep learning before describing what a neural network is. Then, the notion of training a neural network, with all that is involved such as forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation is explained. This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks". They are used in many computer vision applications due to their great performance on these tasks. Finally, the concept of "transfer learning" is discussed as the entire chapter \ref{ch:transfer_learning} relies on it.

@@

Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, translators, computer-aided diagnosis systems, personal assistants, art creation, robotics, etc. The presence of deep learning techniques in these use cases clearly confirm the enthusiasm of many for this technology. Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=X Axis, 
          ylabel=Y Axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {1/(1+exp(-x))};
        \addlegendentry{$g(x)$}
        \addplot+[no marks] {(1/(1+exp(-x))) * (1-(1/(1+exp(-x))))};
        \addlegendentry{$g(x)(1-g(x))$}
      \end{axis}
    \end{tikzpicture}
    \caption{The sigmoid function and its derivative}
  \end{center}
\end{figure}

@@

  \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=X Axis, 
          ylabel=Y Axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {(1-exp(-2*x))/(1+exp(-2*x))};
        \addlegendentry{$tanh(x)$}
        \addplot+[no marks] {1-((1-exp(-2*x))/(1+exp(-2*x)))^2};
        \addlegendentry{$1-tanh^{2}(x)$}
      \end{axis}
    \end{tikzpicture}
    \caption{The tanh function and its derivative}
  \end{center}
\end{figure}

@@

 \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}[
    declare function={
    func(\x)= (\x<=0) * (0);
  }
  ]
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=X Axis, 
          ylabel=Y Axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {(x>=0)*x};
        \addlegendentry{$ReLU(x)$}
        \addplot[domain=-20:0,red] {0};
		\addplot[domain=0:20, red] {1};
        \addlegendentry{$ReLU'(x)$}
      \end{axis}
    \end{tikzpicture}
    \caption{The ReLU function and its derivative (undefined when $x=0$)}
  \end{center}
\end{figure}

@@

  \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}[
    declare function={
    func(\x)= (\x>=0) * (x) +and (\x<0) * 4;
  }
    ]
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=X Axis, 
          ylabel=Y Axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot[domain=0:20, blue] {x};
        \addlegendentry{$ELU(x)$}
        \addplot[domain=0:20,red] {1};
        \addlegendentry{$ELU'(x)$}

@@

In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting the parameters of the model. The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network. The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ at point $x$ is given by:
\begin{equation}
\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{m}}]
\end{equation}
The gradient shows how much the parameters that constitute x need to change to minimize the function. In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$. The computation of all these gradients relies on the "chain rule". In the case of weights, the chain rule is:
\begin{equation}
\frac{\partial J}{\partial w_{jk}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial w_{jk}^{l}}
\end{equation}
Similarly, the chain rules has to be applied to the biases:
\begin{equation}
\frac{\partial J}{\partial b_{j}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial b_{j}^{l}}
\end{equation}
Once the gradients of each parameter are computed, the corresponding parameters are updated. The weight update is described by the following equation:
\begin{equation}
W^{[l]} = W^{[l]} - \alpha * \frac{\partial J}{\partial W^{[l]}}
\end{equation}
The bias update corresponds to:
\begin{equation}
b^{[l]} = b^{[l]} - \alpha * \frac{\partial J}{\partial b^{[l]}}
\end{equation}
%The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be manually tuned.

@@

The simplest and least efficient technique to initialize neural network weights is to randomly generate them. The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values, which lead to value near 0 or 1 for most activation functions. Consequently, the slope of the gradient changes slowly and the learning process takes a lot of time.\\
To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
\begin{equation}
\frac{\sqrt{6}}{\sqrt{n_{h}^{[l]}+n_{h}^{[l+1]}}}
\end{equation}
where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
For activation functions that are not symmetric around zero and do not have outputs inside the range $[-1,1]$ such as ReLU or ELU, Kaiming initialization is an alternative. It consists in multiplying the randomly initialized weight matrix by:
\begin{equation}
\frac{\sqrt{2}}{\sqrt{n_{h}^{[l]}}}
\end{equation} 
where $n_{h}^{[l]}$ is the number of incoming connections coming to a given layer from the previous layer's output.

@@

Finally, the number of epochs during which the network is trained has to be carefully chosen. In fact, from a certain point in the training onwards, neural networks do not learn anymore useful features in the data and start overfitting. This point corresponds to the moment where the validation loss does not decrease anymore and starts to increase continuously. It is exactly at this moment that the training should stop. To achieve this goal, it is possible either to directly choose the right number of epochs or to use the so-called "early stopping" method, which stops the training as soon as the validation loss does not decrease for $N$ epochs.

@@

To perform a convolution, numerical input data and a "filter" are required. A filter can be seen as an $f$x$f$ numerical patch that moves across the entire input. It first moves horizontally until reaching the right-most border of the image. Then, it goes down a cell and starts from the border on the left-hand side. This process is repeated until the filter reaches the bottom-right corner of the image, which marks the end of the convolution. At each step, the dot product between the filter and the part of the input covered by the filter is computed. Figure \ref{fig:convolution} showcases a simple convolution which aims at finding vertical lines in a black and white image. For example, the output of the first step of the convolution (in red) is computed by evaluating the dot product between the blue filter and the red part of the input image: 
\begin{equation}
\begin{gathered}
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}0
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}= -2
\end{gathered}
\end{equation}

\noindent The output of the entire convolution shows negative values in the outside parts and large positive values in the center. This means that the $3x3$ filter detected a vertical line in the center of the input image. Like other parameters, filters are learned while training.

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution.png}
\caption{Convolutions - Basic convolution in a CNN}
\label{fig:convolution}
\end{figure}
\noindent Different parameters can change the way a convolution behaves. First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom. In this case, the so-called "stride" is equal to 1. Other applications could rely on a bigger stride. Furthermore, the previous example reduced the output size of $3$x$3$ in proportion to the initial input size of $5$x$5$. To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s. Three ways of padding images are commonly used as shown on Figure \ref{fig:convolution_padding}:
\begin{itemize}
	\item \textbf{Valid:} The input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.

@@

\section{Transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task"~\cite{30}. The model dedicated to the second task uses some or all parts of the first model (i.e. keeps the same weights and architecture or a part of them) and is then retrained on data available for this task. The first model can either be implemented from scratch if enough data is available or downloaded from institutions that release large pretrained models for similar tasks. Transfer learning provides three major benefits~\cite{30}:
\begin{itemize}
\item Higher start: The initial skill (before refining the model) on the source model is higher than it otherwise would be.
\item Higher slope: The rate of skill improvement during training of the source model is steeper than it otherwise would be.
\item Higher asymptote: The converged skill of the trained model is better than it otherwise would be.
\end{itemize}
Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated"~\cite{30}.

>>>>

\chapter{Deep learning}
\label{ch:deep_learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}This chapter provides the theoretical foundation in deep learning which is required to understand the rest of the work. It starts with a historical timeline of deep learning before describing what a neural network is. Then, the notion of training a neural network, with all that is involved (like forward propagation, backpropagation, hyperparameters, data splitting or performance evaluation), is explained. This part is followed by another section devoted to a special type of neural networks, the "convolutional neural networks". They are used in many computer vision applications due to their great performance on these tasks. Finally, the concept of "transfer learning" is discussed as Chapter \ref{ch:transfer_learning} entirely relies on it.

@@

Neural networks are inspired from the biological neural networks of the brain. These systems try to learn how to solve a problem based on the data they receive as input. Many concrete applications make use of neural networks: autonomous vehicles, translators, computer-aided diagnosis systems, personal assistants, art creation, robotics, etc. The presence of deep learning techniques in these use cases clearly confirms the enthusiasm of many for this technology. Furthermore, as this field has recently gained interest (see Section \ref{historical_background}), a lot of research is still ongoing, which suggests that many exciting new applications will certainly be discovered in the near future.

@@

\begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=x-axis, 
          ylabel=y-axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {1/(1+exp(-x))};
        \addlegendentry{$g(x)$}
        \addplot+[no marks] {(1/(1+exp(-x))) * (1-(1/(1+exp(-x))))};
        \addlegendentry{$g(x)(1-g(x))$}
      \end{axis}
    \end{tikzpicture}
    \caption{The sigmoid function and its derivative}
  \end{center}
\end{figure}

@@

  \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=x-axis, 
          ylabel=y-axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {(1-exp(-2*x))/(1+exp(-2*x))};
        \addlegendentry{$tanh(x)$}
        \addplot+[no marks] {1-((1-exp(-2*x))/(1+exp(-2*x)))^2};
        \addlegendentry{$1-tanh^{2}(x)$}
      \end{axis}
    \end{tikzpicture}
    \caption{The tanh function and its derivative}
  \end{center}
\end{figure}

@@

 \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}[
    declare function={
    func(\x)= (\x<=0) * (0);
  }
  ]
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=x-axis, 
          ylabel=y-axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot+[no marks] {(x>=0)*x};
        \addlegendentry{$ReLU(x)$}
        \addplot[domain=-20:0,red] {0};
		\addplot[domain=0:20, red] {1};
        \addlegendentry{$ReLU'(x)$}
      \end{axis}
    \end{tikzpicture}
    \caption{The ReLU function and its derivative (undefined when $x=0$)}
  \end{center}
\end{figure}

@@

  \begin{figure}[h!]
  \begin{center}
    \begin{tikzpicture}[
    declare function={
    func(\x)= (\x>=0) * (x) +and (\x<0) * 4;
  }
    ]
      \begin{axis}[
      	  width=0.55\linewidth, % Scale the plot to 0.7 \linewidth
          xlabel={$x$}, 
          ylabel={$y$},
          xlabel=x-axis, 
          ylabel=y-axis,
          samples=41, 
          grid, 
          thick,
          domain=-10:10,
		legend style={at={(0,1)},anchor=north west}
        ]
        \addplot[domain=0:20, blue] {x};
        \addlegendentry{$ELU(x)$}
        \addplot[domain=0:20,red] {1};
        \addlegendentry{$ELU'(x)$}

@@

In other words, after each forward pass through the entire network, backpropagation performs a backward pass which aims at minimizing the cost function by adjusting the parameters of the model. The way parameters are updated is defined by the gradients of the cost function with respect to each parameter of the network. The gradient of the cost function $J(x_{1}, x_{2}, ..., x_{m})$ at point $x$ is given by:
\begin{equation}
\frac{\partial J}{\partial x} = [\frac{\partial J}{\partial x_{1}}, \frac{\partial J}{\partial x_{2}}, ..., \frac{\partial J}{\partial x_{m}}]
\end{equation}
The gradient shows how much the parameters that constitute $x$ need to change to minimize the function. In neural networks, the parameters of the cost function are all weight matrices $W^{[l]}$ and biases $b^{[l]}$. The computation of all these gradients relies on the "chain rule". In the case of weights, the chain rule is:
\begin{equation}
\frac{\partial J}{\partial w_{jk}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial w_{jk}^{l}}
\end{equation}
Similarly, the chain rules has to be applied to the biases:
\begin{equation}
\frac{\partial J}{\partial b_{j}^{l}} = \frac{\partial J}{\partial z_{j}^{l}} * \frac{\partial z_{j}^{l}}{\partial b_{j}^{l}}
\end{equation}
Once the gradients of each parameter are computed, the corresponding parameters are updated. The weight update is described by the following equation:
\begin{equation}
W^{[l]} = W^{[l]} - \alpha * \frac{\partial J}{\partial W^{[l]}}
\end{equation}
The bias update corresponds to:
\begin{equation}
b^{[l]} = b^{[l]} - \alpha * \frac{\partial J}{\partial b^{[l]}}
\end{equation}
%The hyperparameter $\alpha$ of these two equations is called "learning rate". It determines the gradient's influence at each epoch and has to be manually tuned.
The "learning rate" $\alpha$ determines the influence that the gradient has at each epoch. It is an hyperparameter and has to be manually tuned.

@@

The simplest and least efficient technique to initialize neural network weights is to randomly generate them. The major problem of this technique comes from the fact that some initializations can lead to extremely small or big values that lead to values near 0 or 1 for most activation functions. Consequently, the slope of the gradient changes slowly and the learning process takes a lot of time.\\
To prevent this effect when the tanh activation function is used, "Xavier initialization" multiplies the random initialization by the fraction:
\begin{equation}
\frac{\sqrt{6}}{\sqrt{n_{h}^{[l]}+n_{h}^{[l+1]}}}
\end{equation}
where $n_{h}^{[l]}$ is the number of incoming network connections to the layer and $n_{h}^{[l+1]}$ is the number of outgoing network connections from that layer.\\
For activation functions that are not symmetric around zero and do not have outputs inside the range $[-1,1]$ such as ReLU or ELU, Kaiming initialization is an alternative. It consists in multiplying the randomly initialized weight matrix by:
\begin{equation}
\frac{\sqrt{2}}{\sqrt{n_{h}^{[l]}}}
\end{equation} 
where $n_{h}^{[l]}$ is the number of incoming connections coming to a given layer from the previous layer's output.

@@

Finally, the number of epochs during which the network is trained has to be carefully chosen. In fact, from a certain point in the training onwards, neural networks do not learn anymore useful features in the data and start overfitting. This point corresponds to the moment where the validation loss does not decrease anymore and starts increasing continuously. It is at this moment that the training should stop. To achieve this goal, it is either possible to directly choose the right number of epochs or to use the so-called "early stopping" method which stops the training as soon as the validation loss does not decrease for $N$ epochs.

@@

To perform a convolution, numerical input data and a "filter" are required. A filter can be seen as an $f*f$ numerical patch that moves across the entire input. It first moves horizontally until reaching the right-most border of the image. Then, it goes down a cell and starts from the border on the left-hand side. This process is repeated until the filter reaches the bottom-right corner of the image, which marks the end of the convolution. At each step, the dot product between the filter and the part of the input covered by the filter is computed. Figure \ref{fig:convolution} showcases a simple convolution which aims at finding vertical lines in a black and white image. For example, the output of the first step of the convolution (in red) is computed by evaluating the dot product between the blue filter and the red part of the input image: 
\begin{equation}
\begin{gathered}
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}0
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}+ \\ 
\color{blue}(-1) \color{black}* \color{red}0 \color{black}+ \color{blue}2 \color{black}* \color{red}0 \color{black}+ \color{blue}(-1) \color{black}* \color{red}1 
\\ \color{black}= -2
\end{gathered}
\end{equation}

\noindent The output of the entire convolution shows negative values in the outside parts and large positive values in the center. This means that the $3*3$ filter detected a vertical line in the center of the input image. Like other parameters, filters are learned while training.

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/convolution.png}
\caption{Convolutions - Basic convolution in a CNN}
\label{fig:convolution}
\end{figure}
\noindent Different parameters can change the way a convolution behaves. First of all, the previous example relied on a filter moving by respectively one cell to the right and to the bottom. In this case, the so-called "stride" is equal to 1. Other applications could rely on a bigger stride. Furthermore, the previous example reduced the output size of $3*3$ in proportion to the initial input size of $5*5$. To influence the output size, a padding can be added to the outside of the input image, usually filled with 0s. Three ways of padding images are commonly used as shown on Figure \ref{fig:convolution_padding}:
\begin{itemize}
	\item \textbf{Valid:} The input image is not padded. This means that the filter only goes through existing pixel values, which makes the output size smaller than the input size.

@@

\section{Transfer learning}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin}}According to Jason Brownlee, "transfer learning is a machine learning method where a model developed for a task is reused as the starting point for a model on a second task"~\cite{30}. The model dedicated to the second task uses some or all parts of the first model (i.e. keeps the same weights and architecture or a part of them) and is then retrained on the data that is available for this task. The first model can either be implemented from scratch if enough data is available or downloaded from institutions that release large pretrained models for similar tasks. Transfer learning provides three major benefits~\cite{30}:
\begin{itemize}
\item Higher start: The initial skill (before refining the model) on the source model is higher than it otherwise would be.
\item Higher slope: The rate of skill improvement during training of the source model is steeper than it otherwise would be.
\item Higher asymptote: The converged skill of the trained model is better than it otherwise would be.
\end{itemize}
Nevertheless, "in general, it is not obvious that there will be a benefit to using transfer learning in the domain until after the model has been developed and evaluated"~\cite{30}.

### 2020-03-14 Common/chapters/literature_review.tex
Julien and Johan: minors

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training set ($80\%$), validation set ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal and vertical flipping, horizontal and vertical shifting of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
Their model is a modified version of the well-known VGG16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer. They also use the ELU activation function instead of ReLU (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the augmentation of the test set. The best results were obtained by using DWI images with the highest b-value (stacked with T2 and ADC images), reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge, which would have been an interesting benchmark to evaluate how well the model generalizes.

@@

\section{Lung/Lung CT Challenge}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Gao et al. claimed that "lung cancer is one of the most frequent and leading causes of death all over the world. It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The papers cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans, and are devoted to the classification of lung nodules (see Section \ref{sec:lungCTChallenge} for further details).

Cengil et al.~\cite{02} built a fairly simple convolutional neural network to classify the images of the Lung CT Challenge dataset. The model takes 4-dimensional data as input (depth, height, width and channels) and performs 3D convolutions on it. The model consists of an input layer, five layers of 3D convolutions (the first is associated with a RELU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end. Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.

@@

\section{Brain/Kaggle Brain}

@@

Saxena et al.~\cite{31} implemented three convolution neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. Their processing method used a cropping technique which removed extra black margin around the skull. Each border of the image was merged with a part of the skull. Since the images come from different sources, their resolution vary quite a lot. Therefore, the authors resized them to $224$x$224$x$3$. Moreover, data was augmented by rotation, vertical shifting and horizontal shifting. As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting. The data was split into a training set, a validation set and a test set. Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance. The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$). The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 didn't perform well (AUC of $0.55$ and accuracy \mbox{of $0.55$}).

>>>>

Song et al.~\cite{07} presented a DCNN method to detect prostate cancer on multiparametric MRIs. Their data processing approach kept T2W, DWI and ADC grayscale images only. After resampling each image to the same resolution, T2W, DWI and ADC images were first cropped ($65$x$65$px patch) with the lesion in the center and stacked per patient, resulting in images containing three grayscale channels. Thanks to this method, the same lesion is visible in the same area over the three channels. This increases the probability of detecting a cancer by ensuring a good visibilty for each lesion, since the latter is not necesarily as visible with each parameter. Images were then normalized based on the Z-score per patient and per sequence (T2W, DWI, ADC), i.e. by subtracting the mean before dividing by the standard deviation. The data was split into a training set ($80\%$), validation set ($10\%$) and test set ($10\%$). The training (undefined number of times), validation (undefined number of times) and test images (11x) were augmented using -20$^\circ$ to 20$^\circ$ rotations, horizontal and vertical flipping, horizontal and vertical shifting of less than 2 pixels and stretching by a factor between 0.9 and 1.1.
Their model is a modified version of the well-known VGG-

16 model, including the addition of $1$x$1$ convolutions and dropout layers after each max pooling layer. They also use the ELU activation function instead of ReLU (see figure \ref{fig:paper_model}). The evaluation method for each patient and finding made an average of the 11 predictions resulting from the augmentation of the test set. The best results were obtained by using DWI images with the highest b-value (stacked with T2 and ADC images), reaching an AUC of $0.944$ with a 95$\%$ confidence interval ($0.876$-$0.994$). However, this model was not tested on the official PROSTATEx challenge, which would have been an interesting benchmark to evaluate how well the model generalizes.

@@

\section{Lung -- Lung CT Challenge}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Jobin \& Cl{\'e}ment}}Gao et al. claimed that "lung cancer is one of the most frequent and leading causes of death all over the world. It was reported that there were approximately $1.8$x$10^6$ new cases of lung cancer globally in 2012. Early detection of lung cancer, which is typically viewed in the form of lung nodules, is an efficient way to improve the survival rate"~\cite{41}. The papers cited below make use of the SPIE-AAPM Lung CT Challenge dataset, which is composed of CT scans, and are devoted to the classification of lung nodules (see Section \ref{sec:lungCTChallenge} for further details).

Cengil et al.~\cite{02} built a fairly simple convolutional neural network to classify the images of the Lung CT Challenge dataset. The model takes 4-dimensional data as input (depth, height, width and channels) and performs 3D convolutions on it. The model consists of an input layer, five layers of 3D convolutions (the first is associated with a ReLU activation function and pooling, the last with nothing, and the others with pooling) and a fully connected layer at the end. Regarding the model evaluation, authors announce an accuracy of 0.7 on their test set which contains 30 findings.

@@

\section{Brain -- Kaggle Brain}

@@

Saxena et al.~\cite{31} implemented three convolutional neural networks to classify the brain tumors coming from the Kaggle "Brain MRI Images for Brain Tumor Detection" dataset. Their processing method used a cropping technique which removed extra black margin around the skull. Each border of the image was merged with a part of the skull. Since the images come from different sources, their resolution varies quite a lot. Therefore, the authors resized them to $224$x$224$x$3$. Moreover, data was augmented using rotation, vertical shifting and horizontal shifting. As the cropping was performed before augmenting the images, small parts of the brain were outside of the augmented images due to rotation and shifting. The data was split into a training set, a validation set and a test set. Regarding the models, authors implemented three of them (a Resnet-50, a VGG-16 and an Inception-V3) in order to compare their performance. The best results on the test set were achieved by the Resnet-50 (AUC of $0.95$ and accuracy of $0.95$). The VGG-16 was close (AUC of $0.90$ and accuracy of $0.90$), whereas the Inception-V3 did not perform well (AUC of $0.55$ and accuracy \mbox{of $0.55$}).
